The good, the bad, and the ugly

### 2025-02-05 paper/algorithms.tex
cose

\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black]
\tikzstyle{decision} = [diamond, minimum width=1.5cm, minimum height=0.5cm, text centered, draw=black]
\tikzstyle{arrow} = [thick,->,>=stealth]

@@

\maketitle

\begin{abstract}

\end{abstract}

@@

\section{Bruto}

$ \theta = \{ x \in \mathbb{R} \,|\, \theta_{min} \leq x \leq \theta_{max} \}$: Latent trait levels, where $\theta_{min}$ is the minimum value of $\theta$ and $\theta_{max}$ is its maximum value \add{(non sono sicura della x)}

$||\theta||$: cardinality of the latent trait

$\forall Q \in\mathcal{Q} = 2^B \setminus \{\emptyset, B\}$,

\begin{enumerate}
	\item $\mathbf{TIF}^{Q} =  \frac{\sum_{i \in Q} IIF_i}{||Q||}$
	\item $\overline{\Delta}_{\mathbf{TIF}^{Q}} =  \mathit{mean}(|\mathbf{TIF}^* - \mathbf{TIF}^{Q}|)$  \add{ho dei dubbi su questa notazione. la media è calcolata attraverso i valori di theta}. In alternativa a questa notazione: $\overline{\Delta}_{\mathbf{TIF}^{Q}} = \frac{|\mathbf{TIF}^* - \mathbf{TIF}^{Q}|}{||\theta||}$

\end{enumerate}

$Q_{bruto} = \arg \min_{Q \in \mathcal{Q}} \overline{\Delta}_{\mathbf{TIF}^{Q}}$

\section{Item Locating Algorithm}

\add{Siccome la filosofia di ILA e ISA è molto simile sarebbe carino trovare loro un nome comune e poi declinarle nelle loro specificità. ci pensiamo}

\textbf{Set up: }

$B$: Item bank

$Q^k \subset B$: Set of item indexes selected for inclusion in the STF up to iteration $k$

$\mathbf{TIF}^*$: TIF target

$D$: Item selected at each iteration

$\mathbf{PIF}_D^k = \frac{\sum_{i\in Q^k \cup \{D^k\}} IIF_i}{||Q^k \cup \{D\}||}$, provisional mean tif at iteration $k$, including the item selected in $D^k$

$||Q^k||$: cardinality of $Q^k$ at iteration $k$

At $k = 0$: $\mathbf{PIF}_D^0 = (0, 0, \ldots, 0)$, $Q^0 = \emptyset$, $D^0 = \emptyset$,

\begin{enumerate}
	\item $\theta_{target} := \arg \max |\mathbf{TIF}^* - \mathbf{PIF}_D^{k}|$
	\item $D^k = \arg \min_{i \in B\setminus Q^k} |\theta_{target} - b_i|$
	\item $\mathbf{PIF}_D^k = \frac{\sum_{i \in Q^k \cup \{D\}} IIF_i}{||Q^k \cup \{D\}||}$
	\item Termination Criterion: $|\mathbf{TIF}^* - \mathbf{PIF}_D^k| < |\mathbf{TIF}^* - \mathbf{TIF}^{k}|$: 
	\begin{itemize}
		\item TRUE: $k := k +1$, $Q^{k} = Q^{k-1} \cup \{D\}$, iterates 1-5 (\add{Ho scritto $k-1$ perché siccome ho scritto all'inizio che $k$ si aggiorna e diventa $k+1$ mi sembrava sensato})
		\item FALSE: The item in $D$ does not contribute to reduce the distance from the TIF target, hence: $Q_{ISA} = Q^k$
	\end{itemize}
\end{enumerate}

\section{Item Selecting Algorithm}

Same as ILA but based on the Item Information Functions.

\color{blue}

Set up same as ILA: 
$B$: Item bank

$Q^k \subset B$: Set of item indexes selected for inclusion in the STF up to iteration $k$

$\mathbf{TIF}^*$: TIF target

$D$: Item selected at each iteration

$\theta_{target}$: theta target selected at each iteration

$\mathbf{PIF}_D^k = \frac{\sum_{i\in Q^k \cup \{D\}} IIF_i}{||Q^k \cup \{D\}||}$, provisional mean tif at iteration $k$, including the item selected in $D$

$||Q^k \cup \{D\}||$: cardinality of $Q^k \cup \{D\}$ at iteration $k$

@@

At $k = 0$: $\mathbf{PIF}_D^0 = (0, 0, \ldots, 0)$, $Q^0 = \emptyset$, $D^0 = \emptyset$,

\begin{enumerate}
	\item $\theta_{target} := \arg \max |\mathbf{TIF}^* - \mathbf{PIF}_D^{k}|$
	\item $IIF_{i \in B \setminus Q^k} = a_i^2P(\theta_{target}, a_i, b_i)[1-P(\theta_{target}, a_i, b_i)]$
	\item $D := \arg \min_{i \in B\setminus Q^k} |\theta_{target} - IIF_i|$
	\item $\mathbf{PIF}_{D^k} = \frac{\sum_{i \in Q^k \cup \{D\}} IIF_i}{||Q^k \cup \{D\}||}$
	\item Termination Criterion: $|\mathbf{TIF}^* - \mathbf{PIF}_D^k| < |\mathbf{TIF}^* - \mathbf{TIF}^{k}|$: 
	\begin{itemize}
		\item TRUE: $k := k +1$, $Q^{k} = Q^{k-1} \cup \{D\}$, iterates 1-5 (\add{Ho scritto $k-1$ perché siccome ho scritto all'inizio che $k$ si aggiorna e diventa $k+1$ mi sembrava sensato})
		\item FALSE: The item in $D$ does not contribute to reduce the distance from the TIF target, hence: $Q_{ISA} = Q^k$
	\end{itemize}
\end{enumerate}

\section{Frank}

The setup is like the one of ILA and ISA:

$B$: Item bank

$Q^k$: set of items selected at iteration $k$

$D$: provisional item selected at each iteration

\add{$\mathbf{PIF}$: provisional mean tif}

At $k =0$, $\mathbf{PIF}^0 = (0, 0, \ldots, 0)$, $Q^0 = \emptyset$, iterate

\begin{enumerate}
	\item  $A^k = B \setminus Q^k$ (sets of available items at iteration $k$)
	\item $\forall i \in A^k$, $\mathbf{PIF}_{i}^k = \frac{\mathbf{TIF}^k + \mathbf{IIF}_{i}}{||Q^k||+1}$
	\item $D = \arg \min_{i \in A^k} |\mathbf{TIF}^* - \mathbf{PIF}_i^k|$
	\item Termination criterion: $|\mathbf{TIF}^* - \mathbf{PIF}_D^{k}| < |\mathbf{TIF}^* - \mathbf{TIF}^{k-1}|$: 
	\begin{itemize}
		\item  If true, $k := k + 1,$ $Q^{k} = Q^{k-1} \cup \{D\}$, restart from 1 \add{C'è il meno 1 per la stessa ragione scritta per ILA e ISA}

		\item  If false, stops, $Q_{Frank} = Q^k$

>>>>

\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
\tikzstyle{arrow} = [thick,->,>=stealth]

% formattazione per APA 7

\raggedbottom
\titleformat*{\section}{\bfseries\centering}
\titleformat*{\subsection}{\bfseries\flushleft}
\titleformat*{\subsubsection}{\bfseries\itshape\flushleft}
\titleformat*{\paragraph}{\bfseries}
\titleformat*{\subparagraph}{\large\bfseries}

@@

\vspace*{10mm}
\begin{center}
	\begin{LARGE}
	The Good, the Bad, and the Ugly: \\ New Item Response Theory based algorithms for the development of short test forms
	\end{LARGE}

	\vspace{5mm}
	\begin{large}
%		Ottavia M. Epifania, Pasquale Anselmi, and Egidio Robusto \\ 
%		Department of Philosophy, Sociology, Education, and Applied Psychology, University of Padova
	\end{large}
\end{center}
\vspace{15mm}
\begin{center}
	\textbf{Author Note}
\end{center}
\begin{small}

%	Ottavia M. Epifania \includegraphics[width=0.06\linewidth]{orcid.png} https://orcid.org/0000-0001-8552-568X
%	
%	Pasquale Anselmi \includegraphics[width=0.06\linewidth]{orcid.png} https://orcid.org/0000-0003-2982-7178 
%	
%	Egidio Robusto \includegraphics[width=0.06\linewidth]{orcid.png} https://orcid.org/0000-0002-7583-2587

	The authors have no conflict of interest to disclose.

	Correspondence concerning this article should be addressed to: Ottavia M. Epifania, Department of Philosophy, Sociology, Education, and Applied Psychology, Via Venezia 14, Padova, Italy.
	E-Mail: ottavia.epifania@unipd.it \\
\end{small}

\vfill

%\end{document}

\begin{abstract}
	\onehalfspacing
This manuscript presents Item Response Theory-based algorithms for the development of short test forms (STFs) from full-length tests. All algorithms aim at minimizing both the number of items included in the STF and the minimum average distance from a test information target that describes the desired characteristics of the STF, denoted as test information function target (TIF-target). The algorithms differ according to the method used for the item inclusion in the STF at each iteration, while they share the same termination criterion. 
The item selection is based on the ability of each item to reduce the distance from the TIF-target considering either one specific latent trait level defined at each iteration or the entire latent trait level.
The algorithms stop when the last selected item does not contribute in reducing the distance from the TIF-target. 
The performance of the algorithms is compared against that of a brute force algorithm, which selects the item combination best able to reduce the distance from the TIF-target among all the possible item combinations.  Results of a simulation study show that the algorithm that selects the items according to their ability of reducing the distance from the TIF-target throughout the entire latent trait outperforms the other algorithms. Limitations of the study and future lines of research are discussed.

\textbf{Keywords:} Item Response Theory; Information Functions; Algorithms; Short Test Forms
\end{abstract}

@@

\newpage
\begin{center}
	\textbf{The Good, the Bad, and the Ugly: New Item Response Theory-Based Algorithms for the Development of Short Test Forms}
\end{center}

%Since Item Response Theory (IRT) provides detailed information on the measurement precision of each item across different levels of the latent trait, it constitutes an ideal framework for test development \cite<e.g.,>{baker2017, van1987absolute}. Moreover, it allows for the development of tests or short test forms  (STFs) from an item bank or from an existing test, respectively.
Item Response Theory (IRT) offers detailed insights into the measurement precision of individual items across varying levels of the latent trait, making it an ideal framework for test development \cite<e.g.,>{baker2017, van1987absolute}. By leveraging its robust analytical capabilities, IRT facilitates the construction of tailored assessments, such as tests designed from item banks or the derivation of short test forms (STFs) from existing assessments. This flexibility ensures that tests can be optimized for specific purposes, populations, or contexts while maintaining precision and validity.
The STFs developed from tests can be classified into two main types.
 On the one hand, there are the adaptive STFs developed within the computerized adaptive testing framework, which tailor the item administration to maximize measurement precision for each respondent \cite<e.g.,>{magis}. Therefore, the items may differ from one respondent to another. On the other hand, static STFs are composed of the same subset of items, which is chosen to maximize the measurement precision across all the administered respondents.   
%The information on the measurement precision of each item can be exploited for the creation of ad-hoc STFs by selecting only those items that most precisely measure specific levels of the latent trait, in line with the aim of the assessment.  
This article presents three new IRT-based algorithms for the development of STFs. 
All algorithms attempt at recreating the desired characteristics of a test, as defined through a target information function, by iteratively selecting the items that reduce as much as possible the distance between the information function of the STF and the target one. In this application, the focus is on the development of static STFs from full-length tests, although the introduced algorithms also apply to the development of static tests from item banks.
%The methods used for selecting the items at each iteration differentiate the algorithms. Specifically, they select the items according to the item's ability of reducing the distance from the TIF-target considering either a single latent trait level defined at each iteration or the entire latent trait.

In common IRT procedures \cite<e.g.,>{chiesi, colledani2018, silvia}, the items  are selected for inclusion in the STF according to their measurement precision with respect to different levels of the latent trait that are of interest for the assessment. To pursue this aim, the information functions of all the items of the full-length test are inspected, and the items with the information functions best able to comply with the desired characteristics of the STF are selected for inclusion. 
Although this procedure offers high degrees of freedom (e.g., the number of items in the STF can be not specified in advance), it strongly relies on the expertise and subjectivity of the researchers, which lead to issues related to 
replicability.   
Moreover, since this procedure is not automated, it might be particularly demanding, especially if the full-length test comprises many items. 
Recently, \citeA{pauci} introduced a procedure, denoted as $\theta$-target procedure, which integrates the principles of computerized adaptive testing (i.e., tailoring the item administration to specific latent trait levels) for the development of static STFs from full-length tests, thus allowing for its automation. 
The newly introduced procedure grounds the item selection on the measurement precision of each item in the full-length test with respect to specific latent trait levels of interest, denoted as $\theta$ targets, which are defined a priori according to the aim of the assessment. 
The procedure selects an optimal item (i.e., the item that provides the highest measurement precision) for each of the identified $\theta$ targets. 
Despite being promising, the discrete definition of the $\theta$ targets raises different issues that cannot be ignored. 
Firstly, since an optimal item is selected for each of the $\theta$ targets, the actual number of $\theta$ targets determines the number of items included in the STFs. 
As such, the more the STF is meant to precisely measure larger regions of the latent trait, the larger the number of selected $\theta$ targets and the larger the number of items that need to be included in the STF. 
Moreover, the functioning of the $\theta$-target procedure strongly depends on the availability in the item bank of informative items with respect to the defined $\theta$ targets. If the item bank does not contain informative items with respect to one or more $\theta$ targets, sub-optimal items might be included in the final item selection. As such, the STF might contain more items than actually needed or items that might not appropriately assess the levels of interest. 
Finally,  the $\theta$ targets are defined as discrete values to recreate a sort of ``overall'' information function of the STF, which is continuous by nature. 
%The $\theta_{target}$s are defined a priori according to the specific aim with which the STF is developed. For instance, if the aim of the STF is to discriminate respondents below and above a certain cut-off (e.g., clinical diagnostic tests) the $\theta$s are defined around the cut-off point, such that only the most informative items for that region of the latent trait are included in the STF. Otherwise, if the aim is to maximize the information across the entire latent trait, the $\theta_{target}$s can be equally spread along the latent trait continuum.   
%Typically, the development of STF relies on the expertise of the researchers, in that that the items are manually chosen according to the measurement precision that they provide to address the final measurement precision one wants to obtain for the STF. While such a procedure provides large degrees of freedom in terms of item selection and does not require the definition of the number of items a priori, it strongly relies on the expertise of the researcher and might be at risk of not being reproducible. 
%While both types of STFs are based on the same principle, they differ in their focus. Adaptive STFs adjust to the individual respondent’s characteristics, maximizing measurement precision for that specific person. In contrast, static STFs aim to achieve overall precision across the entire group of respondents. Both approaches leverage the information from IRT models regarding the measurement precision of each item at various levels of the latent trait.
In this light, it might be more convenient to directly refer to a continuous function describing the desired characteristics of the STF by means of a target information function to be recreated with the final selection of items in the STF \cite<e.g.,>{lordTif}.
The specification of a TIF-target allows for shifting the focus on different regions of the latent trait, hence STFs addressing different aims can be developed. 
\begin{figure}
	\centering
	\begin{subfigure}{0.45\linewidth}
		\includegraphics[width=\textwidth]{img/target_n}
		\caption{}
		\label{fig:targetN}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\includegraphics[width=\textwidth]{img/target_sk}
		\caption{}
		\label{fig:targetSK}
	\end{subfigure}
\caption{Examples of target information functions. Panel \ref{fig:targetN} illustrates a target information function aimed at obtaining a good measurement precision of latent trait levels between $-2$ and $2$. Panel \ref{fig:targetSK} illustrates a target information function focused on a specific level of the latent trait, $\theta = -2$.}
	\label{fig:tartge-examples}
\end{figure}
For instance, Figure \ref{fig:tartge-examples} illustrates two TIF-targets, which have been specified for obtaining either  precise measurements of different levels of the latent trait. 
Figure \ref{fig:targetN} illustrates a TIF target for obtaining a precise measurement of the $\theta$ levels between $-2$ and $2$. Figure \ref{fig:targetSK} illustrates a TIF target highly informative for $\theta$ levels around $-2$ standard deviation below the average population mean of 0 \cite<a cut off at 2 standard deviations below the population mean is a criterion sometimes used in diagnostic tests, see e.g., >{balboni, tasse2016}. 
%Specifically, Figure \ref{fig:targetN} depicts the TIF-target that allows for obtaining precise measurements across almost the entire latent trait, with a particular focus on the regions between $-2$ and $2$. Such a TIF-target might be employed when developing STFs for the assessment of psychological constructs that are normally distributed among the general population (e.g., intelligence). On the other hand, Figure \ref{fig:targetSK} depicts a TIF-target that is consistent for the development of a diagnostic STF, since its focus is on a specific $\theta$ level

\normalcolor
Although the specification of the desired characteristics of the STF through a target information function is reasonable, its actual definition might not be straightforward. While there might be an idea concerning the shape of the target information function (i.e., the regions of the latent trait on which the highest measurement precision is desired), the definition of its height (i.e., the amount of information for different levels of the latent trait) might be troublesome \cite{van1989relative}. 
In this sense, if both the shape and height of the target information function can be defined, the test or STF development is performed with the idea of recreating an \emph{absolute information} \cite{stocking, van1987absolute}. 
Conversely, if only the shape of the target information function is known and can be defined, the test or STF development is performed with the idea of recreating the shape of a \emph{relative information} \cite{van1989relative}.

%\subsection{Existing Procedures}

%Adaptive STFs are . According to this procedure, the administration of each item depends on the responses registered at the previous items. 
%At each iteration, a temporary estimate of the latent trait level of the respondent is computed based on the responses registered up to that point, and the next administered item is the one that can maximize the information for that level.
%As such, each respondent can be administered with a different subset of items, which aims at maximizing the measurement precision for each respondent.

%The latent trait levels on which the static STFs focus the most might change and should be defined in advance. For instance, if the STF is aimed ad the assessment of clinical population with medium-low latent trait levels, the items composing the STF should be mostly located in those regions. The STF may not be optimal in terms of maximizing measurement precision for each individual respondent. However, it would still ensure adequate measurement precision for respondents within those specific latent trait levels, providing reliable measurement for those groups.

%Although adaptive STFs are undoubtedly convenient since they allow for obtaining precise estimates of the latent trait level of each respondent by tailoring the item administration, they might be problematic in some contexts. For instance, they require the implementation of adaptive algorithms on ad-hoc platforms for the administration, which might not be always accessible or they might be difficult to use with some target populations (e.g., the elderly). Moreover, fairness issues might be raised in evaluation contexts where it is important to administered all respondents with the same subset of items.

This manuscript presents three new IRT-based algorithms for the development of STFs, which ground the item selection on the definition of the general desired characteristics of the STF as expressed by a target test information function. The continuous definition of the desired characteristics allows for overcoming the issues related to the a priori specification of $\theta$ targets, hence releasing the constraint on the optimal number of items. As the  target information function is empirically defined with a specific shape and height, these algorithms attempt at finding the item selection to include in the STF that is best able to reduce the distance from an absolute information.
All algorithms consider the distance between the target information function and the information function obtained on the items selected up to that iteration with the aim of bridging the gap between the two information functions (i.e., bringing the information function of the STF as close as possible to the target one). As it will be further illustrated, they differ according to how they consider the distance between the information functions and hence on how they consider the items for inclusion in the STF. The performance of the algorithms is evaluated in a simulation study, where it is compared against that of a brute force algorithm. 
Since the brute force algorithm selects the item combination best able to reduce the distance from the target information function among all the possible item combinations of different length, its item selection is considered as the gold standard.

@@

% This contribution is focused on the development of STFs from full-length tests. Nonetheless, the presented algorithms can be also used for the development of informative tests developed from item banks.

The manuscript is organized as follows. The next section illustrates the IRT model used as reference in this work, along with the definition of the related information functions. The proposed algorithms are then presented, followed by a simulation study that compares their performance in STFs development. The manuscript concludes with final remarks.

%Come limitazione il fatto che il numero dei theta target influenza molto il numero di item da inserire, tanto che più theta target ci sono più item devono essere includi nella forma breve ma forse più importante il fatto che è basata su valori puntuali del tratto latente quando forse sarebbe più comodo definer una tif generale da riprodurre (riprendere il discorso di Tif relativa e Tif assoluta e dire chiaramente che in questo caso ci muoviamo nel caso di una tif assoluta visto che viene definita un'altezza dell'infromazione ma che in futuro ci contreeremo sulla tif relativa)

\section*{Item Response Theory and Information Functions}
Different IRT models are available for the analyses of dichotomous responses, which differentiate according to the characteristics of the items that they take into account. 
In this application, 
the 2-Parameter Logistic model \cite<2-PL;>{birnbaum1968} is used as a reference, which considers item difficulty and item discrimination. According to this model, the probability of observing a positive response by person $p$ on item $i$ is formalized as:

\begin{equation}
	P(x_{pi} = 1|\theta_p, b_i, a_i) = \dfrac{\exp [a_i(\theta_p - b_i)]}{1 + \exp [a_i(\theta_p - b_i)]},
\end{equation}
where $\theta_p$ is the ability parameter of person $p$ (i.e., the latent trait level of $p$), $b_i$ is the difficulty parameter of item $i$ (i.e., the location of $i$ on the latent trait), and $a_i$ is the discrimination parameter of item $i$ (i.e., the ability of $i$ to distinguish respondents with different levels of $\theta$). The probability of observing a positive response changes as the distance between the ability of the person $\theta_p$ and the location of the item on the latent trait $b_i$ changes, such as it is of the 50\% when the ability of person $p$ matches the difficulty of item $i$.

The precision with which each item measures different levels of the latent trait is described by the \emph{item information function}, $\text{IIF}_i = a_i^2P(\theta, b_i, a_i)[1- P(\theta, b_i, a_i)]$. In the 2-PL, the IIF reaches its maximum when $\theta_p = b_i$. The precision with which the test as a whole measures the entire latent trait is expressed by the \emph{test information function}, which is the sum of the IIFs of all the items in a test or item bank, $\text{TIF} = \sum_{i \in B} \text{IIF}_i$ (where $B$ is the set of items in a test or in an item bank). 
Given the dependency between the location of the items on the latent trait, their discrimination parameters, and their informativeness with respect to different latent trait levels,  the shape and height of the TIF depends on the spread of the items along the latent trait continuum.
 Since the TIF is the sum of the IIFs, the higher the number of items in  test, the higher the TIF. Alternatively, a mean TIF can be computed by averaging for the number of items. As it will be further illustrated, the computation of the mean TIF might be more convenient when developing STFs.

\section*{The Algorithms}

The three introduced algorithms aim at approximating as best as possible the desired characteristics of a test, as expressed by a target information function (TIF-target), by selecting the items best able to reduce the distance between the information function of the STF and the target one. 
Two of the algorithms (item locating algorithm -- ILA -- and item selecting algorithm -- ISA --) can be seen as the natural evolution of the $\theta$-target procedure in that they ground the item selection on the characteristics of each item with respect to a specific $\theta$ target, which is defined at each iteration as the latent trait level where the maximum distance between the two information functions is observed.
The third one (Frank) overcomes the definition of the $\theta$ targets altogether by considering the entire latent trait for the item inclusion at each iteration. As such, it accounts for the measurement precision of each item with respect to the entire latent trait and to the TIF-target. 
All three algorithms attempt at bridging the gap between the TIF-target and the temporary information function obtained from the items selected up to that iteration. 
By grounding the item selection on a specific $\theta$ target, ILA and ISA bring the two information functions closer considering only one latent trait level at the time. 
On the other hand, by grounding the item selection on its contribution to the temporary information function, Frank bridges the gap between the two functions considering the entire latent trait. 
%However, ILA and ISA attempt at bridging the gap between the two information functions considering one discrete level at the time, the $\theta$ target, while Frank considers the entire latent trait. 
%As such, ILA and ISA select the best item according to one specific point of the latent trait, while Frank  selects the item that allows for bridging the gap between the information functions throughout the entire latent trait. 
%\color{red} Forse questa parte si può togliere 
%In ILA and ISA, the $\theta$ targets are defined at each iteration as the level of the latent trait for which the maximum distance between the information function obtained from the items selected up to that iteration and the target information function is observed. 
%Differently from the $\theta$-target procedure, the number of items does not need to be defined a priori and the algorithms automatically searches for the levels of the latent trait that need to be addressed at each iteration. The main issue of ILA and ISA pertains the items selection grounded on discrete levels of the latent trait, instead of considering its continuous nature. 
%Frank aims at overcoming this issue by considering the continuous nature of the target information function. At each iteration, it explores the distance between the target information function and the information function obtained with the items selected up to that iteration, and selects the item that is best able to bridge the gap between the target and the temporary information function considering the entire latent trait. 
\normalcolor

The ability of ILA, ISA, and Frank to approximate an absolute TIF-target is evaluated by comparing their performance against that of a brute force algorithm.
This algorithm
 tries every possible combination of items without repetition and selects the one best able to bridge the gap between the two information functions among all the possible item combinations attainable from the full-length test. As such, its resulting STF is considered as the gold standard.

All algorithms but the brute force start with an empty item collection, $Q$, in which an item is included at each iteration until the termination criterion is met.  %Since each algorithm starts with an empty set and adds one item at each iteration, the mean TIF is computed at each iteration by considering the number of items in the current set $Q$.
Before any item can be included in $Q$, the termination criterion tests its usefulness in reducing the distance from the TIF-target by concurrently considering the TIF-target, the TIF obtained from the STF with the last item considered for inclusion (i.e., provisional item), and the TIF obtained from the STF without this last item. 
%At each iteration, the termination criterion tests whether other items should be added in the STF by considering the distance between the TIF-target, the TIF obtained by adding the last item considered for inclusion (i.e., provisional item), and the TIF obtained without this item. 
If the distance between the former two is greater than or equal to the distance between the latter two, the provisional item is not included in the STF and the algorithm stops. Vice versa, the item is included and the algorithms proceeds to a new iteration. The reasoning is as follows. If the distance between the TIF-target and the STF including the provisional item is greater than or equal to the difference between the TIF-target and the STF without the provisional item, the provisional item does not contribute in reducing this distance. 
Since the provisional item is the best one among the available options according to the selection method of each algorithm and does not contribute to reduce the distance from the TIF-target, the algorithm stops. 
When the termination criterion is reached, the final item selection included in the STF is the one in $Q$ up to the last iteration, which does not include the provisional item.

In what follows, the notation $||X||$ is used to denote the cardinality of a generic set $X$, $B$ denotes the set of items in the full-length test, $Q_x$ indicates the subset $Q_x \subset B$ of items included in the STF by each algorithm $x \in \{\text{Bruto}, \text{ILA}, \text{ISA}, \text{Frank}\}$, and $\text{TIF}^*$ indicates the TIF-target. The TIF-target is used to express the desired characteristics of the STF in terms of absolute information across the latent trait. 
Since the focus of the manuscript is on the development of STFs from full-length tests, $Q_x$ is taken to be a proper subset of the items in the test, such that it is not possible to obtain a STF that contains all the items of the test. If the algorithms are applied for the generation of informative tests from an item bank, $Q_x$ can be the subset of $B$, and all the items can be included in the test.

%\add{current number of items in the STF, va ben definito da qualche parte. nei cofnronit si parla di una stf che è in divenire a partire  da un insieme vuoto di item e che la media è calcolata sulla base del numero di item che sono presenti in quel momento anella STF}

The brute force algorithm (i.e., Bruto) does not have a proper termination criterion given that it generates all the possible STFs of different length and then chooses the one that allows for minimizing the distance from the $\text{TIF}^*$. It follows that Bruto does not have iterations either. 
%For all other algorithms, the termination criterion is the same. 
%Specifically, it tests whether the last provisional item (i.e., the item that is being evaluated for the inclusion in the STF at the current iteration) is useful in reducing the distance from the $\text{TIF}^*$. 
%If the distance between the $\text{TIF}^*$ and the provisional TIF that includes the last provisional item (denoted as $\text{pTIF}$) is greater than or equal to the distance between the $\text{TIF}^*$ and the TIF obtained without the provisional item, this is deemed as not useful in reducing the distance from the $\text{TIF}^*$ and the algorithm stops.  The subset of items would hence be the one at the previous iteration, excluding the provisional item. 
%If the same comparison is false, the provisional item is deemed as useful in reducing the distance from the $\text{TIF}^*$, it is included in the subset of items and a new iteration starts. 
%Not having a termination criterion, Bruto does not have iterations either. Indeed, Bruto generates all the possible STFs of different lengths considering different combination of items, and it chooses the one that is best able to reduce the distance from the $\text{TIF}^*$.
%Bruto is the only algorithm for which the number of iterations is known in advance, since it compares the TIF attainable from all the possible combinations of items of different length from an item bank \add{non so se sia corretto parlare di number of iterations}. 
The number of iterations performed by the other algorithms cannot be known in advance since they iteratively compare the TIF obtained by including the provisional item with the $\text{TIF}^*$ until the termination criterion is met. 
The only constraints on the number of iterations pertains the lower and upper bound. 
 Specifically, since all the algorithms start from an empty set of items, they are forced to perform at least two iterations such that at least one item can be included in the STF. Moreover, the upper constraint is logically imposed by the number of items in the full-length test, such that the maximum number of iterations is equal to the cardinality of the full-length test. Importantly, if the algorithms reach the last iteration and the termination criterion is not met, it is considered as a failure in the their ability to find a STF able to reproduce the $\text{TIF}^*$.

 Finally, since the TIF is usually obtained by summing the IIFs together, the more the items in a test, the higher the TIF. 
 As such, when the TIF is used as criterion for deciding the best subset of items to include in a STF, the risk is to favor STFs that include more items. 
 To avoid this risk, the mean TIF of the STF is used in the comparison against the TIF-target. In what follows, the mean  TIF is simply indicated as TIF.

\subsection*{Bruto}

Bruto compares the TIF of all the possible item combinations attainable from the items in $B$ against the $\text{TIF}^*$ and selects the item combination best able to reduce the average distance from the $\text{TIF}^*$, as follows:

%\begin{tikzpicture}[node distance=2cm]
%	
%	% Nodes
%	\node (start) [startstop] {Start: $\forall Q \in\mathcal{Q} = 2^B \setminus \{\emptyset, B\}$};
%	\node (step1) [process, below of=start] {Step 1: Compute $\mathbf{TIF}^Q = \frac{\sum_{i \in Q} IIF_i}{||Q||}$};
%	\node (step2) [process, below of=step1] {Step 2: Compute $\overline{\Delta}_{\mathbf{TIF}^Q} = \mathit{mean}(|\mathbf{TIF}^* - \mathbf{TIF}^Q|)$};
%	\node (end) [startstop, below of=step2] {End: $Q_{bruto} = \arg \min_{Q \in \mathcal{Q}} \overline{\Delta}_{\mathbf{TIF}^Q}$};
%	
%	% Arrows
%	\draw [arrow] (start) -- (step1);
%	\draw [arrow] (step1) -- (step2);
%	\draw [arrow] (step2) -- (end);
%	
%\end{tikzpicture}

$\forall Q \in\mathcal{Q} = 2^B \setminus \{\emptyset, B\}$, compute

\begin{enumerate}
	\item $TIF^{Q} =  \frac{\sum_{i \in Q} IIF_i}{||Q||}$ and
	\item $\overline{\Delta}_{\text{TIF}^{Q}} =  \mathit{mean}(|TIF^* - TIF^{Q}|)$. 
\end{enumerate}

Then, $Q_{bruto} = \arg \min_{Q \in \mathcal{Q}} \overline{\Delta}_{\mathbf{TIF}^{Q}}$.

For all the possible item combinations in $\mathcal{Q}$ that are attainable from the item bank $B$ (excluding the empty set and the full set of items), Bruto computes: (1.) the mean TIF, $TIF^Q$ and (2.) 
the average distance $\overline{\Delta}_{\text{TIF}^Q}$ between the $\text{TIF}^*$ and the $\text{TIF}^Q$.  Then, the final subset of items of the STF, $Q_{bruto}$, is the one that allows for obtaining the least value of $\overline{\Delta}_{\text{TIF}^{Q}}$, hence for minimizing the distance from the $\text{TIF}^*$.

Given that Bruto attempts all the possible item combinations in $\mathcal{Q}$, it ensures to find the one that is best able to minimize the average distance from the $\text{TIF}^*$.
% that aims at the selection of best combination of items to recreate a given $TIF^*$  by comparing the TIF-target with the TIF obtained with all the possible combinations of items of different lengths that are obtainable from an item bank $B$. The total number of item combinations of different lengths $\mathcal{Q}$ is equal to the power set of the item bank minus the empty set and the full set of items. For each item combination $Q \in \mathcal{Q}$, the mean TIF $TIF^Q$ is computed, and is compared against the TIF-target. Specifically, the average distance between the TIF-target and $TIF^Q$ is computed and the item combination selected by Bruto is the one among all item combinations that allows fro minimizing the distance between $TIF^*$ and $TIF^Q$.

\subsection*{Item Locating Algorithm}

At each iteration $k$, the item locating algorithm (ILA) selects the item that minimizes the distance on the latent trait from a specific $\theta$ target (denoted as $\theta_{target}$), as follows:

At $k = 0$: $\text{TIF}^0(\theta) = 0 \, \forall \theta$, $Q^0 = \emptyset$. For $k \geq 0$,
\begin{enumerate}
	\item $\theta_{target} := \arg \max |\text{TIF}^* - \text{TIF}^{k}|$
	\item $i^* := \arg \min_{i \in B\setminus Q^{k}} |\theta_{target} - b_i|$
	\item $\text{pTIF}_{i^*} = \frac{TIF^k + IIF_{i^*}}{||Q^{k}|| + 1}$ %$\mathbf{pTIF}_{i^*}= \frac{\sum_{i \in Q^{k} \cup \{i^*\}} IIF_i}{||Q^{k} \cup \{i^*\}||}$ $=$ 
	\item Termination Criterion: $|\text{TIF}^* - \text{pTIF}_{i^*}| \geq |\text{TIF}^* - \text{TIF}^{k}|$: 
	\begin{itemize}
		\item FALSE:  $Q^{k+1} = Q^{k} \cup \{i^*\}$, $\text{TIF}^{k+1} = \text{pTIF}_{i^*}$, iterates 1-4 
		\item TRUE: Stop, %The item in $i^*$ does not contribute to reduce the distance from the TIF-target, hence: 
		$Q_{ILA} = Q^k$
	\end{itemize}
\end{enumerate}
